---
title: "Lab 1 - Data visualization"
author: "Jack Morgenstein"
format: pdf
editor: visual
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)
```

## Exercise 1

```{r}
ggplot(midwest, mapping = aes(x = popdensity)) + 
  geom_histogram(binwidth = 10000, alpha = 0.8, color = "black", 
                 fill = "grey") +
  labs(x = "Population Density", y = "Count", title = 
         "Population Densities of Midwest Counties")
```

The data distribution is strongly right skewed. A vast majority of counties have population densities between 0 and 10,000 while a much smaller number of counties have any higher population densities. There appear to be two outlier bins. A small number of counties have population densities between 55 and 65 thousand people as well as between 85 and 95 thousand people.

## Exercise 2

```{r}
ggplot(midwest, mapping = aes(x = percollege, y = percbelowpoverty, 
                              color = state)) +
  geom_point() + 
  labs(x = "Percent of Population with a College Degree", 
       y = "Percentage Living Below Poverty Line", 
       title = "College Education versus Poverty Rates in the Midwest",
       fill = "State") +
  theme_light()
```


## Exercise 3

(Type your answer to Exercise 3 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

## Exercise 4

## Exercise 5

## Exercise 6

## Exercise 7
